Decision Support and Health Insurance Choice

### 2019-11-07 _Abstract_ASHEcon_201910.Rmd
Split cleanup files and update analysis to household level

---
title: "Decision Support and Health Insurance Choice"
author: Ian McCarthy & Evan Saltzman
date: October 2019
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
fontsize: 12pt
subparagraph: yes
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{titlesec}
   - \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
   - \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
   - \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
output: 
  pdf_document:
    keep_tex: TRUE
bibliography: 'D:/CloudStation/Professional/Bibliography/BibTeX_Library.bib'
---

@@

### Data:
We have data on `r format(tot.enroll, big.mark=",")` enrollee/year observations from the California health insurance exhanges from `r first.year` to `r last.year`. Our data include a variety of individual and household characteristics, plan choices and plan characteristics, and information on the type of decision assistance used by the enrollee (if any). From these data, we can identify each household's set of possible health insurance plans, and we employ premium and cost sharing subsidy formulas to calculate health insurance costs for each possible plan for each household.

### Methods:
We first estimate overall demand for health insurance using a nested logit discrete choice model, as in @saltzman2019. We then identify specific instances in which the observed plan choice is dominated by some other plan in an individual's choice set. For example, Gold and Platinum tier plans are dominated for any household that is eligible for cost-sharing subsidies and with incomes below 150% of the federal poverty level. We estimate the effect of decision assistance on dominated choices with a linear probability model, allowing for year and household fixed effects.

### Results:
Our nested logit models establish strong evidence that decision assistance matters for insurance choice. Turning specifically to dominated choices, our baseline results suggest that individuals with decision assistance are `r abs(round(reg2$coeff[1],4))*100` percentage points less likely to make a dominated choice. On a base of `r round(mean.dom.choice,3)*100`%, this reflects a `r abs(round(reg2$coeff[1]/mean.dom.choice,2))*100`% decrease in the probability of making a dominated choice.

We also find strong heterogeneities in the effects of different forms of decision assistance. In particular, we find that individuals using publicly provided decision assistance are `r abs(round(reg3$coeff[2],4))*100` percentage points less likely to select a dominated plan, while individuals using a private insurance agent are `r abs(round(reg3$coeff[1],4))*100` percentage points less likely to select a dominated plan.

>>>>

---
title: "Decision Support and Health Insurance Choice"
author: Ian McCarthy & Evan Saltzman
date: October 2019
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
fontsize: 12pt
subparagraph: yes
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{titlesec}
   - \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
   - \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
   - \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
output: 
  pdf_document:
    keep_tex: TRUE
bibliography: 'C:/Users/immccar/CloudStation/Professional/Bibliography/BibTeX_Library.bib'
---

@@

### Data:
Our final dataset consists of `r format(tot.enroll, big.mark=",")` household/year observations from the California health insurance exhanges from `r first.year` to `r last.year`. Our data include a variety of household characteristics, plan choices and plan characteristics, and information on the type of decision assistance used by the enrollee (if any). From these data, we can identify each household's set of possible health insurance plans, and we employ premium and cost sharing subsidy formulas to calculate health insurance costs for each possible plan for each household.

### Methods:
We first estimate overall demand for health insurance using a nested logit discrete choice model, as in @saltzman2019. We then identify specific instances in which the observed plan choice is dominated by some other plan in an individual's choice set. For example, Gold and Platinum tier plans are dominated for any household that is eligible for cost-sharing subsidies and with incomes below 150% of the federal poverty level. We estimate the effect of decision assistance on dominated choices with a linear probability model, allowing for year, insurer, and region fixed effects.

### Results:
Our nested logit models establish strong evidence that decision assistance matters for insurance choice. Turning specifically to dominated choices, our baseline results suggest that individuals with decision assistance are `r abs(round(reg1$coeff[1],4))*100` percentage points less likely to make a dominated choice. On a base of `r round(mean.dom.choice,3)*100`%, this reflects a `r abs(round(reg1$coeff[1]/mean.dom.choice,2))*100`% decrease in the probability of making a dominated choice.

We also find evidence fo heterogeneities in the effects of different forms of decision assistance. In particular, we find that individuals using publicly provided decision assistance are `r abs(round(reg1.b$coeff[2],4))*100` percentage points less likely to select a dominated plan, while individuals using a private insurance agent are `r abs(round(reg1.b$coeff[1],4))*100` percentage points less likely to select a dominated plan.
